%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Author: Sibi <sibi@psibi.in>
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{article}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\setlength\parskip{\baselineskip}
\begin{document}
\title{Chapter 3 (Section 3.1)}
\author{Sibi}
\date{\today}
\maketitle
\newpage

\section{Problem 1}
Consider the following theorem. (This theorem was proven in the
introduction.)

Theorem. Suppose n is an integer larger than 1 and n is not prime.
Then $2^n - 1$  is not prime.

(a) Identify the hypotheses and conclusion of the theorem. Are the
hypotheses true when n = 6? What does the theorem tell you in this
instance? Is it right?

(b) What can you conclude from the theorem in the case n = 15? Check
directly that this conclusion is correct.

(c) What can you conclude from the theorem in the case n = 11?

Solution (a)

Hypotheses:
\begin{itemize}
\item n is an integer
\item $n > 1$
\item n is not prime
\end{itemize}

Conclusion:
$2^n - 1$ is not prime.

Hypotheses are true when $n = 6$. The theorem in the instance says
that $2^n - 1 = 65$ for $n=6$ is not a prime number. Yes, it is
correct.

Solution (b)

When, $n=15, 2^n - 1 = 32767/$. So, I guess that is not an prime
number. 7 is one of the factors.

Solution (c)

When, $n=11, 2^n - 1 = 2047$. So 2047 is not a prime number. And hey
23 is one of it's factor.

\section{Problem 2}
Consider the following theorem. (The theorem is correct, but we will not
ask you to prove it here.)

Theorem. Suppose that $b^2 > 4ac$. Then the quadratic equation $ax^2 +
bx + c = 0$ has exactly two real solutions.

(a) Identify the hypotheses and conclusion of the theorem.

(b) To give an instance of the theorem, you must specify values for a, b,
and c, but not x. Why?

(c) What can you conclude from the theorem in the case a = 2, b = −5,
c = 3? Check directly that this conclusion is correct.

(d) What can you conclude from the theorem in the case a = 2, b = 4, c
= 3?

Solution (a)

Hypotheses:
\begin{itemize}
\item $b^2 - 4ac$
\end{itemize}

Conclusion: $ax^2 + bx + c$ has two real solutions.

Solutions (b)

Because a, b and c are free variables.

Solution (c)

$a = 2, b = -5, c=3$, Since $25 > 24$, they mush have two real
solutions. And infact they have two real solutions: 1 and 1.5.

Solution (d)

When $a =2, b=4, c=3$, $b^2 > 4ac is equivalent to 16 > 24$, so it
doesn't have two real solutions.

\end{document}

